---
title: "Data Analysis Script of the Pointing Experiment Project 2"
output: html_document
---

```{r setup, message=FALSE, echo=FALSE}

## Clear workspace
rm(list=ls())

## The "pacman" package automatically install missing packages and load them
if (!require("pacman")) install.packages("pacman", repos = 'https://stat.ethz.ch/CRAN/'); 

pacman::p_load(
  # data processing
  tidyverse,  # collection of the tidyverse packages
  stringr,    #   - for string functions
  forcats,    #   - utility functions for working with factor levels
  lubridate,  #   - utility for parsing and performing arithematic on dates 
  tools,      #   - for package development, administration and documentation
  data.table, #   - for data manipulation 
  tibble     #   - to create data frames
)
```

## Data Import
Our data is stored in different folder according to the input mode.
```{r import}
# Create a function that imports all .sd2 files from a directory
import_raw_data_file <- function(dir_name, condition_name) {
  
  file_list <- list.files(path = file.path("02 raw data", dir_name), pattern = "*.sd2", recursive = TRUE, full.names = TRUE)
  
  raw_data <- do.call("rbind", lapply(file_list, function(filename)
  data.frame(read.delim(filename, sep = ","))))
  
  # Overwrite Condition because it was used inconsistently during the experiments
  # Therefore, the correct condition needs to be given as a function parameter
  raw_data$Condition = condition_name
  
  return(raw_data)
}

# import all .sd2 files from all conditions and combine them into a single table
data_raw <- rbind(
  import_raw_data_file("1_trackpad", "trackpad"),
  import_raw_data_file("2_index finger", "index finger"),
  import_raw_data_file("3_thumb", "thumb"),
  import_raw_data_file("4_sway", "sway")
)

# Factorise condition
data_raw$Condition <- as.factor(data_raw$Condition)

```

## Data Clean Up
Before we start with the analysis, we will reduce the dataset to the necessary attributes and calcutate the aggregated values.
```{r clean up}
# Keep only the columns of interest and give them more readable names
data_raw <- data_raw %>% 
  select(
    Participant,
    Condition,
    Amplitude = A,
    Width = W,
    PointingTime = PT.ms.,
    SelectionTime = ST.ms.,
    MovementTime = MT.ms.,
    ErrorRate = ER...,
    Throughput = TP.bps.
  )

# Spread the table to make it more suited for further data plot analysis
data_pointing_time <- data_raw %>%
  select(Participant:Width, PointingTime) %>% 
  spread(key = Condition, value = PointingTime) %>% 
  select(Participant:Width,
         PointingTime_trackpad = trackpad,
         PointingTime_index_finger = 'index finger',
         PointingTime_thumb = thumb,
         PointingTime_sway = sway
  )

data_selection_time <- data_raw %>%
  select(Participant:Width, SelectionTime) %>% 
  spread(key = Condition, value = SelectionTime) %>% 
  select(Participant:Width,
         SelectionTime_trackpad = trackpad,
         SelectionTime_index_finger = 'index finger',
         SelectionTime_thumb = thumb,
         SelectionTime_sway = sway
  )

data_movement_time <- data_raw %>%
  select(Participant:Width, MovementTime) %>% 
  spread(key = Condition, value = MovementTime) %>% 
  select(Participant:Width,
         MovementTime_trackpad = trackpad,
         MovementTime_index_finger = 'index finger',
         MovementTime_thumb = thumb,
         MovementTime_sway = sway
  )

data_error_rate <- data_raw %>%
  select(Participant:Width, ErrorRate) %>% 
  spread(key = Condition, value = ErrorRate) %>% 
  select(Participant:Width,
         ErrorRate_trackpad = trackpad,
         ErrorRate_index_finger = 'index finger',
         ErrorRate_thumb = thumb,
         ErrorRate_sway = sway
  )

data_throughput <- data_raw %>%
  select(Participant:Width, Throughput) %>% 
  spread(key = Condition, value = Throughput) %>% 
  select(Participant:Width,
         Throughput_trackpad = trackpad,
         Throughput_index_finger = 'index finger',
         Throughput_thumb = thumb,
         Throughput_sway = sway
  )
```

## Plots for regular edge condition
*Note:* With "trial" we mean the completion of one pointing exercise containing 10 sequences and every sequence is composed of 6 targets on which the participant has to click.

### Movement time per trial for regular edge condition (boxplot)
Boxplot to summarize the times that participant used to complet one task.
```{r boxplot_for_ movement_time}
# Boxplot of the average time participant needed to finish one pointing task
boxplot_movement_time <- boxplot(data_movement_time[4:7], names = c('Trackpad', 'Index Finger', 'Thumb', 'Sway Mode'), xlab = 'Input Mode', ylab = 'Movement Time in ms')
```

### Average movement time per participant for regular edge condition (lineplot)
Lineplot which displays the average time that each participant needed to complete a task.
```{r lineplot_for_movement_time}
# Analysis of average movent time per input mode trial for each participant
average_movement_time_per_participant <- aggregate(data_movement_time[4:7], list(data_movement_time$Participant), mean) %>%
  rename('Participant' = Group.1, 'Trackpad' = MovementTime_trackpad, 'Index Finger' = MovementTime_index_finger, 'Thumb' = MovementTime_thumb, 'Sway Mode' = MovementTime_sway) %>%
  gather(key = Input_Mode, value = Average_Movement_Time, c('Trackpad', 'Index Finger', 'Thumb', 'Sway Mode'))
```

```{r lineplot_for_movement_time_plot, echo=FALSE}
# Lineplot of the calculated movement time average
ggplot (data = average_movement_time_per_participant, mapping = aes(x = fct_relevel(Input_Mode, "Trackpad", "Index Finger", "Thumb", "Sway Mode"), y = Average_Movement_Time, col = Participant, group = Participant)) +
  geom_point() +
  geom_line() +
  labs(x = 'Input Mode', y = 'Movement Time in ms' )
```

### Error rate per task for regular edge condition (boxplot)
Boxplot to summarize the error rates participant had for each completed task.
```{r boxplot_for_error_rate}
# Boxplot of the average error rate that participants have for one pointing task
boxplot_error_rate <- boxplot(data_error_rate[4:7], names = c('Trackpad', 'Index Finger', 'Thumb', 'Sway Mode'), xlab = 'Input Mode', ylab = 'Error Rate in %')
```

### Average error rate per participant for regular edge condition (lineplot)
Lineplot which displays the average error rate each participant had.
```{r lineplot_for_error_rate}
# Analysis of average error rate per input mode trial for each participant
average_error_rate_per_participant <- aggregate(data_error_rate[4:7], list(data_error_rate$Participant), mean) %>%
  rename('Participant' = Group.1, 'Trackpad' = ErrorRate_trackpad, 'Index Finger' = ErrorRate_index_finger, 'Thumb' = ErrorRate_thumb, 'Sway Mode' = ErrorRate_sway) %>%
  gather(key = Input_Mode, value = Average_Error_Rate, c('Trackpad', 'Index Finger', 'Thumb', 'Sway Mode'))
```
```{r lineplot_for_error_rate_plot, echo=FALSE}
# Lineplot of the calculated error rate average
ggplot(data = average_error_rate_per_participant, mapping = aes(x = fct_relevel(Input_Mode, "Trackpad", "Index Finger", "Thumb", "Sway Mode"), y = Average_Error_Rate, col = Participant, group = Participant)) +
  geom_point() +
  geom_line() +
  labs(x = 'Input Mode', y = 'Error Rate in %')
```

## Plots for virtual edge condition
*Note:* With "trial" we mean the completion of one pointing exercise containing 10 sequences and every sequence is composed of 6 targets on which the participant has to click.

### Movement time per trial for virtual edge condition (boxplot)
Boxplot to summarize the times that participant used to complet one task.
```{r boxplot_for_ movement_time}
# Boxplot of the average time participant needed to finish one pointing task
boxplot_movement_time <- boxplot(data_movement_time[4:7], names = c('Trackpad', 'Index Finger', 'Thumb', 'Sway Mode'), xlab = 'Input Mode', ylab = 'Movement Time in ms')
```

### Average movement time per participant for virtual edge condition (lineplot)
Lineplot which displays the average time that each participant needed to complete a task.
```{r lineplot_for_movement_time}
# Analysis of average movent time per input mode trial for each participant
average_movement_time_per_participant <- aggregate(data_movement_time[4:7], list(data_movement_time$Participant), mean) %>%
  rename('Participant' = Group.1, 'Trackpad' = MovementTime_trackpad, 'Index Finger' = MovementTime_index_finger, 'Thumb' = MovementTime_thumb, 'Sway Mode' = MovementTime_sway) %>%
  gather(key = Input_Mode, value = Average_Movement_Time, c('Trackpad', 'Index Finger', 'Thumb', 'Sway Mode'))
```

```{r lineplot_for_movement_time_plot, echo=FALSE}
# Lineplot of the calculated movement time average
ggplot (data = average_movement_time_per_participant, mapping = aes(x = fct_relevel(Input_Mode, "Trackpad", "Index Finger", "Thumb", "Sway Mode"), y = Average_Movement_Time, col = Participant, group = Participant)) +
  geom_point() +
  geom_line() +
  labs(x = 'Input Mode', y = 'Movement Time in ms' )
```

### Error rate per task for virtual edge condition (boxplot)
Boxplot to summarize the error rates participant had for each completed task.
```{r boxplot_for_error_rate}
# Boxplot of the average error rate that participants have for one pointing task
boxplot_error_rate <- boxplot(data_error_rate[4:7], names = c('Trackpad', 'Index Finger', 'Thumb', 'Sway Mode'), xlab = 'Input Mode', ylab = 'Error Rate in %')
```

### Average error rate per participant for virtual edge condition (lineplot)
Lineplot which displays the average error rate each participant had.
```{r lineplot_for_error_rate}
# Analysis of average error rate per input mode trial for each participant
average_error_rate_per_participant <- aggregate(data_error_rate[4:7], list(data_error_rate$Participant), mean) %>%
  rename('Participant' = Group.1, 'Trackpad' = ErrorRate_trackpad, 'Index Finger' = ErrorRate_index_finger, 'Thumb' = ErrorRate_thumb, 'Sway Mode' = ErrorRate_sway) %>%
  gather(key = Input_Mode, value = Average_Error_Rate, c('Trackpad', 'Index Finger', 'Thumb', 'Sway Mode'))
```
```{r lineplot_for_error_rate_plot, echo=FALSE}
# Lineplot of the calculated error rate average
ggplot(data = average_error_rate_per_participant, mapping = aes(x = fct_relevel(Input_Mode, "Trackpad", "Index Finger", "Thumb", "Sway Mode"), y = Average_Error_Rate, col = Participant, group = Participant)) +
  geom_point() +
  geom_line() +
  labs(x = 'Input Mode', y = 'Error Rate in %')
```


## Statistical analysis

### ANOVA

#### Speed
Summary of the ANOVA test for Speed:
```{r anova_speed, echo=FALSE}
fit_speed <- aov(MovementTime ~ Condition, data_raw)
summary(fit_speed)
m.res <- resid(fit_speed)
plot(data_raw$MovementTime, m.res, ylab="Residuals", xlab="MovementTime")
```

#### Accuracy
Summary of the ANOVA test for Accuracy:
```{r anova_accuracy, echo=FALSE}
fit_accuracy <- aov(ErrorRate ~ Condition, data_raw)
summary(fit_accuracy)
m.res <- resid(fit_accuracy)
plot(data_raw$ErrorRate, m.res, ylab="Residuals", xlab="ErrorRate")
```

#### Throughput
Summary of the ANOVA test for Throughput:
```{r anova_throughput, echo=FALSE}
fit_throughput <- aov(Throughput ~ Condition, data_raw)
summary(fit_throughput)
m.res <- resid(fit_throughput)
plot(data_raw$Throughput, m.res, ylab="Residuals", xlab="Throughput")
```

We understand from these ANOVA tests that not all the means are equals. We still have to search which pairs of conditions. For that purpose, we condcut a Post Hoc Test.

### Tukey Test

#### Speed
Summary of the Tukey test for Speed:
```{r tukeyhsd_speed, echo=FALSE}
TukeyHSD(fit_speed)
par(mar=c(5,10,4,2)+0.1) # Add additional margin to the left for fitting the label size

# This hack is used to set a more specific title than the default one generated by R
par(col.main = "white")
plot(TukeyHSD(fit_speed), las=1)
par(col.main = "black")
title(main = "Speed - 95% family-wise confidence level")
```


#### Accuracy
Summary of the Tukey test for Accuracy:
```{r tukeyhsd_accuracy, echo=FALSE}
TukeyHSD(fit_accuracy)
par(mar=c(5,10,4,2)+0.1) # Add additional margin to the left for fitting the label size

# This hack is used to set a more specific title than the default one generated by R
par(col.main = "white")
plot(TukeyHSD(fit_accuracy), las=1)
par(col.main = "black")
title(main = "Accuracy - 95% family-wise confidence level")
```


#### Throughput
Summary of the Tukey test for Throughput:
```{r tukeyhsd_throughput, echo=FALSE}
TukeyHSD(fit_throughput)
par(mar=c(5,10,4,2)+0.1) # Add additional margin to the left for fitting the label size

# This hack is used to set a more specific title than the default one generated by R
par(col.main = "white")
plot(TukeyHSD(fit_throughput), las=1)
par(col.main = "black")
title(main = "Throughput - 95% family-wise confidence level")
```
